UZH IT \& AI Emissions Analysis\vspace{-1em}

### 2024-12-18 it-emissions.tex
More meaningful filename
Doc improvements: explain model before explaining data, add appropriate caveats about stochastic approach.

\documentclass[11pt]{article}
\usepackage[margin=0.7in]{geometry}
\usepackage[T1]{fontenc}
\usepackage{libertine}
\renewcommand{\sfdefault}{fvs} % Use as sans-serif
\renewcommand{\ttdefault}{fvm} % Use as monospace
\renewcommand{\ttfamily}{\small\fontfamily{\ttdefault}\selectfont} % Slightly smaller monospace font
\newcommand{\assumption}[1]{{\bf #1}}
%\usepackage{helvet} % Helvetica font for sans-serif
%\usepackage{courier} % Courier font for monospace
\usepackage{hyperref}
\usepackage[backend=biber,style=authoryear,autocite=inline]{biblatex}
\usepackage{titlesec}
\usepackage{enumitem}
\usepackage{titling}

@@

\addbibresource{ref.bib}
\title{UZH IT \& AI Emissions Analysis\vspace{-1em}}
\author{Massimiliano Poletto}
\date{\vspace{-1em}29.11.2024\footnote{Last modified: \today}}

@@

@@

\assumption{In modeling overall university IT emissions as well as potential LLM use,
we assume that the entire university population has a laptop and a mobile phone.}

We do not yet have any information about the breakdown of different functions / job roles, so
\assumption{we assume that IT use is uniform across the university}. This assumption should be
revisited.

\subsection{Embodied emissions}
The estimation of embodied emissions introduces additional assumptions and uncertainties into
the model. Our initial project proposal ignored embodied emissions altogether, but that leaves
too big a gap.

We did some research on the estimated embodied emissions of different types of devices.

@@

Most estimates are in the range of 250 - 400 kg CO$_2$e. The most detailed data comes from
Dell's Product Carbon Footprints (\textcite{dellpcf}), which they publish for each of their products.
A sample of 15 desktops has mean supply chain emissions of 289 kg CO$_2$e with $\sigma = 80$ kg.
\assumption{We model desktop supply chain emissions using a truncated normal distribution with those characteristics.}

@@

\textcite{davy2021} reports data for many Dell servers and estimates for different types of EC2 instances.
The former have mean supply chain emissions of 1252 kg CO$_2$e with $\sigma = 330$ kg, the latter
1912 kg with $\sigma = 885$ kg.
\assumption{We model server supply chain emissions using a truncated normal distribution corresponding
to Dell servers.}

@@

\assumption{We model phone supply chain emissions using a truncated normal distribution with mean
50 kg CO$_2$e and $\sigma = 10$ kg.}

@@

For on-premise equipment, we would like to model the Swiss electricity mix (see below) and, ideally,
daily variations in usage and electricity carbon intensity.
Generic manufacturer estimates of lifetime usage-related emissions are therefore unsuitable.
There are surprisingly few sources of data on device power consumption (as opposed to CO$_2$ emissions).
Based on a smattering of data sheets, primarily from Apple, Cisco, and Dell, we model
power draw of different devices while under load as truncated normal distributions with the
following initial parameters:

@@

Initially, we assume a distribution of duty cycles and power draw when idle for each device.
As a next step, we may simulate power draw at different times of day to integrate with
hourly electricity mix data, but errors in those assumptions may swamp the relatively small
hourly variations in Swiss electricity carbon intensity.

@@

At a high level, we model IT emissions primarily bottom-up, based on device counts and
characteristics, and use a Monte Carlo approach to account for uncertainties.

For on-campus data centers, we would like to use end-to-end data on electricity use (electric bills)
to evaluate our model.

For each device instance, we sample from the appropriate distribution of embodied emissions,
power draw, and duty cycles to obtain the devices characteristics.

Combining power draw and duty cycle with the Swiss electricity mix, we compute annual operational
emissions and add them to the embodied emissions scaled by the device type's expected lifetime.

As mentioned, a possible future refinement will be to model the duty cycle as a function of time of day
and incorporate finer-grained electricity mix data.

>>>>

\documentclass[11pt]{article}
\usepackage[margin=0.7in]{geometry}
\usepackage[T1]{fontenc}
\usepackage{libertine}
\renewcommand{\sfdefault}{fvs} % Use as sans-serif
\renewcommand{\ttdefault}{fvm} % Use as monospace
\renewcommand{\ttfamily}{\small\fontfamily{\ttdefault}\selectfont} % Slightly smaller monospace font
\newcommand{\assumption}[1]{{#1}}
%\usepackage{helvet} % Helvetica font for sans-serif
%\usepackage{courier} % Courier font for monospace
\usepackage{hyperref}
\usepackage[backend=biber,style=authoryear,autocite=inline]{biblatex}
\usepackage{titlesec}
\usepackage{enumitem}
\usepackage{titling}

@@

\addbibresource{ref.bib}
\title{UZH IT \& AI Emissions Analysis\vspace{-1em}}
\author{Massimiliano Poletto}
\date{\vspace{-1em}Last modified: 18.12.2024}

@@

At a high level, we model IT emissions primarily bottom-up, based on device counts and
estimates of their characteristics (embodied emissions, power utilization, and duty cycle).

The model can operate in two modes: deterministic, where each device type has a fixed set of
characteristics based on the mean of values we found in the literature, and stochastic, where
we pick device characteristics from truncated normal distributions with the mean and standard
deviation of the literature values. Of course, normality is a strong and probably incorrect
assumption, but it is a simple initial way to model the uncertainty in our estimates.

For each device instance, we pick embodied emissions, power draw, and duty cycles either
deterministically or by sampling from the appropriate distribution.

Combining power draw and duty cycle estimates with ground-truth data about the Swiss electricity mix,
we estimate annual operational emissions and add them to the embodied emissions scaled by the device
type's expected lifetime.

For on-campus data centers, for which presumably there are electricity bills or other measures of
overall energy use, we would like to use that data to evaluate our model.

A possible future refinement will be to model the duty cycle as a function of time of day
and incorporate finer-grained electricity mix data.

@@

\assumption{In modeling overall university IT emissions as well as potential LLM use,
we assume that the entire university population has a laptop and a mobile phone.}
We do not yet have any information about the breakdown of different functions / job roles, so
\assumption{we assume that IT use is uniform across the university}. Both of these assumptions
should be revisited.

\subsection{Embodied emissions}

For several device types, especially end-user devices such as phones and laptops, embodied emissions
comprise a majority of lifetime emissions. However, the exact amount of these emissions varies
widely between different sources. Our approach has been to find several studies and industry references,
and then estimate means and standard deviations from those that seemed most comprehensive and reliable
(for example, Dell's Product Carbon Footprint reports about its servers).

@@

Most estimates are in the range of 250 - 400 kg CO$_2$e. The most detailed data comes from
Dell's Product Carbon Footprints (\textcite{dellpcf}), which they publish for each of their products.
A sample of 15 desktops has mean supply chain emissions of 289 kg CO$_2$e with $\sigma = 80$ kg.
\assumption{We model desktop supply chain emissions using these statistics.}

@@

\textcite{davy2021} reports data for many Dell servers and estimates for different types of EC2 instances.
The former have mean supply chain emissions of 1252 kg CO$_2$e with $\sigma = 330$ kg, the latter
1912 kg with $\sigma = 885$ kg.
\assumption{We model server supply chain emissions using the statistics corresponding to Dell servers.}

@@

\assumption{We model phone supply chain emissions using mean 50 kg CO$_2$e and $\sigma = 10$ kg.}

@@

For on-premise equipment, we would like to model the Swiss electricity mix (see below) and, ideally,
daily variations in usage and electricity carbon intensity.
Generic manufacturer estimates of lifetime usage-related emissions are therefore unsuitable.
There are surprisingly few sources of data on device power consumption (as opposed to CO$_2$ emissions).
Based on a smattering of data sheets, primarily from Apple, Cisco, and Dell, we model
power draw of different devices while under load using the following initial parameters:

@@

We assume a distribution of duty cycles and power draw when idle for each device.
In the future, we may simulate power draw at different times of day to integrate with
hourly electricity mix data, but errors in those assumptions may swamp the relatively small
hourly variations in Swiss electricity carbon intensity.

@@
